The reference model approach in feature selection problems \vspace{.1in}

### 2019-07-12 paper/ref_approach.tex
language improvements and minor edits

\begin{document}
\maketitle
\begin{abstract}
When a reference model is used to guide the variable selection problem, it acts as a noise-filter modelling the data generation mechanism. We show how this translate into higher stability and better selection of variables. We devise a simple reference model approach that can be used on top of any feature selection procedure and use the normal means problem as a benchmark comparing different methods with or without a reference model. We include in our comparisons state-of-the-art complete selection procedures showing improved stability, false discovery rate and sensitivity of the selection, and also Bayesian methods controlling the amount of posterior shrinkage and bias. Furthermore we show in a real-world example the benefits achieved by the reference model through the projection predictive approach compared to the selection via stepwise backward regression.
\end{abstract}

@@

The goal of the variable selection can be the minimal subset of which
provides good predictive performance (or balance sparsity with
predictive performance), or the all variables which have some
predictive information about the target variable.  In the predictive
performance based variable selection, the future data generation
mechanism can be modelled either using the observed data or a
reference model.
% process is going to be based on assumptions regarding the ``true''
% data generation mechanism.  Making assumptions means making models,
% thus we identify two families of feature selection, or more in general
% model reduction, approaches depending on how the data generation
% mechanism is modelled: the data based and the reference model based
% ones.
The data based approaches, such as the Lasso selection
\cite{tibshirani1996regression} and the stepwise backward/forward
regression the observed empirical data distribution is used as a
proxy of the future data usually in combination with cross-validation
or information criteria to take into account the double use of the
data.
%
The reference model based methods approximate the future data
mechanism using the predictive distribution of a reference model,
which can be, for example, the full-encompassing model. The reference
model approach has been used for long time in the Bayesian framework
since seminal work by \citet{paper:reference_lindley}. For more
historical references see review by \citet{vehtari2012survey} and for
most recent methodological development see paper by
\citet{paper:projpred}.

@@

In this paper we examine in detail the benefits of using a reference
model.  A properly designed reference model is able to filter part of
the noise present in the data, thus to provide an improved and more
stable selection process even if the reference model would not be the
true data generating process. The substantial reduction of the variance is more important than small bias due the potential model misspecification.

@@

The results indicate the core reason why the reference model based
methods perform well is the reference model itself, rather than the
specific way of using it. We do recognize that all these methods
diversify how the subset of variables is actually chosen and which
properties it has, thus it remains fundamental to investigate and
develop such different procedures.
% We hope that this work can increase
% the attractiveness on this family of methods and, therefore, it can be
% of interest to a large number of practitioners.
\\
We would like also the reader to note that although we focus on
Bayesian approach, the benefits of the reference model approach can be
obtained also in frequentist settings.

@@

Relying on the idea of the reference model as a noise-filter, a simple reference model approach, which can be used on top of any feature selection procedure, consists in substituting the target observations with a point prediction of the reference model, e.g. posterior predictive mean. Afterwards, any method for feature selection can be used. Note that we do not argue that such reference model approach should be used as a feature selection tool in real applications; there are many other and more sophisticated ones available in the literature. The reason of this simple approach is to have a fair comparison between using a reference model or not with exactly the same selection procedure. Other kinds of comparison, as the one in Section \ref{projection}, work well as motivational examples, but fail in terms of generality, resulting in a not principled comparison since the presence of the reference model is not the only thing differentiating the two procedures. In Section \ref{comparison} we compare some well-known methods from the literature using both such reference model approach and the original data, observing at the end of the selection higher stability and quality of the selection thanks to the reference model.

@@

As a motivational example to the study of the consequences of the use of a reference model in feature selection problems, we show its application in the form of the projection predictive approach, which uses the reference model to project the posterior samples in the submodels. The \texttt{R}-package \texttt{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and, additionally, it provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. 
\\
We now summarise the workflow of the projection predictive approach in the particular case of the draw-by-draw projection \cite[original formulation by][]{paper:original_proj}, following \cite{paper:projpred}. Assume to have $n$ statistical units with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed covariates, the main steps are the following:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of S posterior samples of the vector parameter of the model.
\item Rank the covariates using a heuristics and consider as candidate submodels the nested ones, starting to include the best ranked feature. The submodel are naturally identify by their model size.
\item For each submodel, project the reference model samples as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL}\;[p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\,||\,p(\tilde{y}_{i}|\boldsymbol{\theta}^{s})] \quad s=1,..,S
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the model distribution of the reference model with parameters fixed at the sample value $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariates related to the statistical unit (identified by the subscript $i$ in the variable $\tilde{y}_{i}$), whereas $p(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ the same but for the submodel. The projected samples correspond to the submodel parameter distribution.
\item For each submodel (size), test the predictive performance.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model prediction utility score.
\end{enumerate}

@@

We show the benefits of the reference model in the form of the projection approach (projpred) with respect to the stepwise backward regression (steplm) using the body fat data \citep{johnson1996fitting}. The analysis is based on from Vehtari's R-notebook available at \url{https://avehtari.github.io/modelselection/bodyfat.html}, which is inspired by \cite{paper:bodyfat}. The target variable of interest is the amount of body fat, which is measured with a complex and expensive procedure, and the original covariates are 13 anthropometric measurements (e.g. height, weight), some of which are highly correlated among themselves and thus provide some challenges in the model selection, for a total of 251 observations. The goal is to find the simplest model which is able to describe, that is to predict, sufficiently well the amount of fat. We use the same data again later in Section \ref{real-world-data}.

@@

We can consider the reference model approach as a family of methods, thus providing an answer to the question what the computational burden is, is not possible, because not unique. It is possible to identify a common and inevitable overhead though, that is the cost of fitting the reference model. In the simple reference model approach we have described at the end of Section \ref{reference-model-approach}, it is actually the only added cost. In general, a full Bayesian model is advisable in order to have the best predictive performance, however including all the available features can be often quite computationally demanding. In such a case a first possible speed-up could be using screening or dimensionality reduction techniques, as for example the supervised principal components \citep{bair2006prediction,piironen2018}. However, in some cases this does not always result in an appreciable speed-up, as when there are many predictive and not correlated features and, thus, many of them have to be included in the reference model anyway. Other alternatives could be fixing some parameters with their maximum marginal likelihood value (typically referred as empirical Bayes) or even use a frequentist approach, if predictions are good enough.

@@

Note that now the quantity of interest $\theta_{j}$ stands for $\sqrt{n-3}\,T_{F}(\rho_{j})$. There are different ways to proceed with the selection from the normal means problem, we consider in our comparison both frequentist and Bayesian state-of-the-art methods. From a full-Bayesian perspective, a Bayesian regression model is fitted to the normal means problem typically using a sparsifying prior as in \cite{paper:dirichlet_laplace}, \cite{bhadra2017horseshoe+} and \cite{johnstone2004needles}. In case of continuous priors, as the Dirichlet-Laplace (DL) and the horseshoe+ priors, there is not a canonical way to complete the selection since the posterior probability of an exactly zero signal is always zero; in their experiments \cite{paper:dirichlet_laplace} use k-means clustering over the posterior samples using the DL prior, whereas \cite{bhadra2017horseshoe+} focus the analysis on the obtained shrinkage and the correct estimation of the sparse signals using the horseshoe+ prior. We include in our experiments the DL and the regularised horseshoe prior \citep{paper:rhs}, which has shown in our opinion better scalability with respect to the horseshoe+, comparing the shrinkage and the correct estimation of the sparse signals using or not the reference model approach. The quantity of interests that we consider are the average SSE (sum of square errors) and the average SESE (sum of expected square errors), defined as follows:

@@

\cite{johnstone2004needles} use a prior of spike-and-slab type \citep{paper:spike_slab_mitchell} given by a mixture of a delta distribution in zero and a heavy-tailed distribution; when the latter is a Laplace distribution they show an interesting thresholding property of the median estimator and, thus, present a complete selection procedure using their prior. We include in our comparison such method together with the control of the local false discovery rate \citep{efron2012large, paper:efron} and a simple selection procedure using inclusion probability at 0.9 for posterior credible intervals using the regularised horseshoe prior (ci.90). All these methods provide a complete selection procedures, we compare the result using or not a reference model on top of the procedure evaluating the average false discovery rate (i.e. ratio of the number of non-relevant selected features over the number of selected features) and the average sensitivity (i.e. ratio of the number of relevant selected features over the total number of relevant features). We also provide a comparison of the stability of the selection using the measure proposed by \cite{paper:stability}. Table \ref{tab:comparison} summarises the methods used in the comparison.

@@

\hypertarget{shrinkage-signal}{%
\subsubsection{Shrinkage and signal analysis}\label{shrinkage-signal}}
In the framework of full-Bayes methods we analyse the effect of the reference model using two continuous hierarchical shrinkage priors: the Dirichlet-Laplace \citep{paper:dirichlet_laplace} and the regularised horseshoe prior \citep{paper:rhs}. Although the horseshoe+ prior by \cite{bhadra2017horseshoe+} is a state-of-the-art prior for sparse signal analysis in the context of the normal means problem, we decided to include the regularised horseshoe due to its faster MCMC inference, since our purpose in this paper is not to study the best prior choice, but rather compare and show the benefits of using a reference model when carrying out variable selection. We used an uniform prior on the interval $(1/p,1/2)$ for the Dirichlet hyperparameter of the Dirichlet-Laplace prior, whereas regarding the reguelarised horseshoe prior we followed the indication provided in \cite{paper:rhs}, setting the prior guess of the effective number of parameters equal to 1 to achieve the highest shrinkage. Yet we have not observed any sensibility of the results due to such a choice.
\
Since, as already mentioned, continuous priors do not usually give an automatic selection procedure, in this Section we study the amount of shrinkage and the bias of the posterior median looking at the SESE and the SSE (see definitions \eqref{eq:SSE} and \eqref{eq:SESE}), together with the posterior mean of the shrink factors.

>>>>

\begin{document}
\maketitle
\begin{abstract}
When a reference model is used to guide the variable selection problem, it acts as a noise-filter modelling the data generation mechanism. We show how this translate into higher stability and better selection of variables. We devise a simple reference model approach that can be used on top of any feature selection procedure and use the normal means problem as a benchmark comparing different methods using or not a reference model. We include in our comparisons state-of-the-art complete selection procedures showing improved stability, false discovery rate and sensitivity of the selection, and also Bayesian methods controlling the amount of posterior shrinkage and bias. Furthermore we show in a real-world example the benefits achieved by the reference model through the projection predictive approach compared to the selection via stepwise backward regression.
\end{abstract}

@@

The goal of the variable selection can be the minimal subset which
provides good predictive performance (or balance sparsity with
predictive performance), or the entire collection of variables which have some
predictive information about the target variable.  In the predictive
performance based variable selection, the future data generation
mechanism can be modelled either using the observed data or a
reference model.
% process is going to be based on assumptions regarding the ``true''
% data generation mechanism.  Making assumptions means making models,
% thus we identify two families of feature selection, or more in general
% model reduction, approaches depending on how the data generation
% mechanism is modelled: the data based and the reference model based
% ones.
In the data based approaches, such as the Lasso selection
\citep{tibshirani1996regression} and the stepwise backward/forward
regression, the observed empirical data distribution is used as a
proxy of the future data usually in combination with cross-validation
or information criteria to take into account the double use of the
data.
%
The reference model based methods approximate the future data generation
mechanism using the predictive distribution of a reference model,
which can be, for example, the full-encompassing model. The reference
model approach has been used for long time in the Bayesian framework
since seminal work by \citet{paper:reference_lindley}. For more
historical references see the review by \citet{vehtari2012survey}, and for
most recent methodological development see 
\citet{paper:projpred}.

@@

In this paper we examine in detail the benefits of using a reference
model.  A properly designed reference model is able to filter part of
the noise present in the data, hence to provide an improved and more
stable selection process even if the reference model would not be the
true data generating process. The substantial reduction of the variance is more important than small bias due the potential model misspecification.

@@

The results indicate the core reason why the reference model based
methods perform well is the reference model itself, rather than the
specific way of using it. We do recognise that all these methods
diversify how the subset of variables is actually chosen and which
properties it has, thus it remains fundamental to investigate and
develop such different procedures.
% We hope that this work can increase
% the attractiveness on this family of methods and, therefore, it can be
% of interest to a large number of practitioners.
\\
We would like also the reader to note that although we focus on
Bayesian approach, the benefits of the reference model approach can be
obtained also in frequentist settings.

@@

Relying on the idea of the reference model as a noise-filter, a simple reference model approach, which can be used on top of any feature selection procedure, consists in substituting the target observations with a point prediction of the reference model, e.g. the posterior predictive mean. Afterwards, any method for feature selection can be used. Note that we do not argue that such reference model approach should be used as a feature selection tool in real applications; there are many other and more sophisticated ones available in the literature. The reason of this simple approach is to have a fair comparison between using a reference model or not with exactly the same selection procedure. Other kinds of comparison, as the one in Section \ref{projection}, work well as motivational examples, but fail in terms of generality, resulting in a not principled comparison since the presence of the reference model is not the only thing differentiating the two procedures. In Section \ref{comparison} we compare some well-known methods from the literature using both such reference model approach and the original data, observing at the end of the selection higher stability and quality of the selection thanks to the reference model.

@@

As a motivational example to the study of the consequences of the use of a reference model in feature selection problems, we show its application in the form of the projection predictive approach, which uses the reference model to project the posterior samples in the submodels. The \texttt{R}-package \texttt{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and, additionally, it provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. 
\\
We now summarise the workflow of the projection predictive approach in the particular case of the draw-by-draw projection \cite[original formulation by][]{paper:original_proj}, following \cite{paper:projpred}. Assume to have $n$ statistical units with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed covariates, the main steps are the following:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of S posterior samples of the vector parameter of the model.
\item Rank the covariates using a heuristics and consider as candidate submodels the nested ones, starting to include the best ranked feature. The submodel are naturally identified by their model size.
\item For each submodel, project the reference model samples as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta^{s}}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL}\;[p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\,||\,p(\tilde{y}_{i}|\boldsymbol{\theta}^{s})] \quad s=1,..,S
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the model distribution of the reference model with parameters fixed at the sample value $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariates related to the statistical unit (identified by the subscript $i$ in the variable $\tilde{y}_{i}$), whereas $p(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ the same but for the submodel. The projected samples correspond to the submodel parameter distribution.
\item For each submodel (size), test the predictive performance.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model prediction utility score.
\end{enumerate}

@@

We show the benefits of the reference model in the form of the projection approach (projpred) with respect to the stepwise backward regression (steplm) using the body fat data \citep{johnson1996fitting}. The analysis is based on Vehtari's \texttt{R}-notebook available at \url{https://avehtari.github.io/modelselection/bodyfat.html}, which is inspired by \cite{paper:bodyfat}. The target variable of interest is the amount of body fat, which is measured with a complex and expensive procedure, and the original covariates are 13 anthropometric measurements (e.g. height, weight), some of which are highly correlated among themselves and thus provide some challenges in the model selection, for a total of 251 observations. The goal is to find the simplest model which is able to describe, that is to predict, sufficiently well the amount of fat. We use the same data again later in Section \ref{real-world-data}.

@@

We can consider the reference model approach as a family of methods, thus providing an answer to the question what the computational burden is, is not possible, because not unique. It is possible to identify a common and inevitable overhead though, that is the cost of fitting the reference model. In the simple reference model approach we have described at the end of Section \ref{reference-model-approach}, it is actually the only added cost. In general, a full Bayesian model is advisable in order to have the best predictive performance, however including all the available features can be often quite computationally demanding. In such a case a first possible speed-up could be using screening or dimensionality reduction techniques, as for example the supervised principal components \citep{bair2006prediction,piironen2018}. However, in some cases this does not always result in an appreciable speed-up, as when there are several predictive and not correlated features and, thus, many of them have to be included in the reference model anyway. Other alternatives could be fixing some parameters with their maximum marginal likelihood value (typically referred as empirical Bayes) or use a frequentist approach, if predictions are good enough.

@@

Note that now the quantity of interest $\theta_{j}$ stands for $\sqrt{n-3}\,T_{F}(\rho_{j})$. There are different ways to proceed with the selection from the normal means problem, we consider in our comparison both frequentist and Bayesian state-of-the-art methods. From a full-Bayesian perspective, a Bayesian regression model is fitted to the normal means problem typically using a sparsifying prior as in \cite{paper:dirichlet_laplace}, \cite{bhadra2017horseshoe+} and \cite{johnstone2004needles}. In case of continuous priors, as the Dirichlet-Laplace (DL) and the horseshoe+ priors, there is not a canonical way to complete the selection since the posterior probability of an exactly zero signal is always zero; in their experiments \cite{paper:dirichlet_laplace} use k-means clustering over the posterior samples using the DL prior, whereas \cite{bhadra2017horseshoe+} focus the analysis on the obtained shrinkage and the correct estimation of the sparse signals using the horseshoe+ prior. We include in our experiments the DL prior and the regularised horseshoe (RHS) prior \citep{paper:rhs}, which has shown in our opinion better scalability with respect to the horseshoe+, comparing the shrinkage and the correct estimation of the sparse signals using or not the reference model approach. The quantity of interests that we consider are the average SSE (sum of square errors) and the average SESE (sum of expected square errors), defined as follows:

@@

\cite{johnstone2004needles} use a prior of spike-and-slab type \citep{paper:spike_slab_mitchell} given by a mixture of a delta distribution in zero and a heavy-tailed distribution; when the latter is a Laplace distribution they show an interesting thresholding property of the median estimator and, thus, present a complete selection procedure using their prior. We include in our comparison such method (EB.med) together with the control of the local false discovery rate \cite[loc.fdr; see][]{efron2012large, paper:efron} and a simple selection procedure using inclusion probability at 0.9 for posterior credible intervals using the regularised horseshoe prior (ci.90). All these methods provide a complete selection procedures, we compare the result using or not a reference model on top of the procedure evaluating the average false discovery rate (i.e. ratio of the number of non-relevant selected features over the number of selected features) and the average sensitivity (i.e. ratio of the number of relevant selected features over the total number of relevant features). We also provide a comparison of the stability of the selection using the measure proposed by \cite{paper:stability}. Table \ref{tab:comparison} summarises the methods used in the comparison.

@@

\hypertarget{shrinkage-signal}{%
\subsubsection{Shrinkage and signal analysis}\label{shrinkage-signal}}
In the framework of full-Bayes methods we analyse the effect of the reference model using two continuous hierarchical shrinkage priors: the Dirichlet-Laplace \citep{paper:dirichlet_laplace} and the regularised horseshoe prior \citep{paper:rhs}. 
% Although the horseshoe+ prior by \cite{bhadra2017horseshoe+} is a state-of-the-art prior for sparse signal analysis in the context of the normal means problem, we decided to include the regularised horseshoe due to its faster MCMC inference, since our purpose in this paper is not to study the best prior choice, but rather compare and show the benefits of using a reference model when carrying out variable selection.
We used an uniform prior on the interval $(1/p,1/2)$ for the Dirichlet hyperparameter of the Dirichlet-Laplace prior, whereas regarding the reguelarised horseshoe prior we followed the indication provided in \cite{paper:rhs}, setting the prior guess of the effective number of parameters equal to 1 to achieve the highest shrinkage. Yet we have not observed any sensibility of the results due to such a choice.
\
Since, as already mentioned, continuous priors do not usually give an automatic selection procedure, in this Section we study the amount of shrinkage and the bias of the posterior median looking at the SESE and the SSE (see definitions \eqref{eq:SSE} and \eqref{eq:SESE}), together with the posterior mean of the shrink factors.
